---
title: "Data Visualization"
author: "Lbrandwo"
date: '2022-04-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## General Information

If we need to be explicit about where a function (or dataset) comes from, we'll use the special form package::function(). For example, `ggplot2::ggplot()` tells you explicitly that we're using the ggplot() function from the ggplot2 package.

Here we analyse the `ggplot2::mpg` dataset which is a data frame of variables (in the columns) and observations (in the rows).mpg contains observations collected by the US Environmental Protection Agency on 38 models of car.

## Creating a ggplot

`displ` is a cars engine size, in litres. `hwy` is a cars fuel efficiency on the highway in miles per gallon.

We would not use a scatterplot for class and drv because they are both categorical.

```{r mpg}
#ggplot(data = <DATA>) + 
        #<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy))

```

## Aesthetics Mappings

We can map colours to variables to indicate more information about them. Here we map `colour` to `class` and we see that the cars with relatively good mpg despite having a large engine are in fact 2seater sports cars.

```{r finding outliers}
ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

Using alpha for discrete variables is not very clear. Furthermore, using shape for more than 6 discrete variables in not advised as it becomes hard to distinguish between them.

```{r shape and alpha}
# Left
ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

For changing aesthetic properties of the graph in general we can just specify `colour` outside the `aes()`.

```{r changing aes}
ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

We can set many `aes()` for each display and even change `colour` based on conditionals.

```{r}
ggplot(data = mpg) +
        geom_point(mapping = aes(x = displ, y = hwy, size = class, shape = trans, stroke = 0.5, colour = hwy < 20))
```

## Facets

Facets are subplots that each display one subset of the data. `facet_wrap()` takes a categorical variable like `class` following `~`, and the `nrow` which you want to spread the facets.

```{r facet_wrap}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

We can facet a plot with a combination of 2 variables using `facet_grid()` which takes 2 categorical variables seperated by `~`. Using `.` as in, `+ facet_grid(. ~ cyl)`, won't facet in the rows or columns dimension. Where you put the `.` determines the axis of facetation.

```{r facet_grid}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

## Different geoms

With `geom_smooth()` we can assign `linetype` to a categorical variable to draw a diff line for each. Here we can overlay 2 geoms and pass the global mappings to ggplot and then specificy plot specific aes individually as well. Instructions given withint the geom functions override global settings for that plot only. `filter()` allows selection of specific data from `mpg`.

```{r multiple geoms}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)     
```

We can choose to not show the legend by assigning the `group` aesthetic to `drv`. Or turning off the legend itself with `show.legend = FALSE`.

```{r turning off legends}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

## Statistical Transformations

*bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

*smoothers fit a model to your data and then plot predictions from the model.

*boxplots compute a robust summary of the distribution and then display a specially formatted box.

Bar plots can be by count or by proportion and we can show multiple variables within each bar which can also be organized by proportion.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")

ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

`position = "jitter"` adds some random noise to scatterplots which separates the points a little for easier visualization of overlapping values.`geom_jitter()` is a short version of `geom_point(position = "jitter")`.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")

```

## Coordinate Systems

* `coord_flip()` switches the x and y axes. This is useful (for example), if you want horizontal boxplots. It’s also useful for long labels: it’s hard to get them to fit without overlapping on the x-axis.

```{r coord flip}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

* `coord_quickmap()` sets the aspect ratio correctly for maps. This is very important if you’re plotting spatial data with ggplot2 (which unfortunately we don’t have the space to cover in this book).

```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

* `coord_polar()` uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```
This is how you flip a stacked bar plot.

```{r}
ggplot(data = diamonds) +
        geom_bar(mapping = aes(x = cut, fill = clarity, position = "fill")) +
        coord_polar()
        
```
* `geom_abline()` adds the reference line and `coord_fixed()` fixes the x and y axis to represent the same movement in each direction per tick.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

